---
title: "Lab #1"
output: html_document
date: "2023-08-30"
author: "Dylan Maddux and Taylor Kingery"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### **LAB 1**

### **Question 1**
```{r}
my_data <- read.csv("https://remiller1450.github.io/data/HappyPlanet.csv")
```

### **Question 2**
```{r}
J <- seq(from = 0, to = 100, by = 10)
K <- rep("XYZ", times = 11)

my_df <- data.frame(J = J, K = K)
```

### **Question 3**
```{r}
str(my_data)
```
There are 11 observations in this data set and 7 variables are numeric

### **Question 4**
```{r}
my_data [57,3]
```

### **Question 5 P1**
```{r}
lastThree <- my_data [141:143,]
lastThree
```

### **Question 5P2**
```{r}
lastTen <- my_data [134:143,]
lastTen
median(lastTen$LifeExpectancy)
```

### **Question 6**:
```{r}
my_data[my_data$Country == "Albania",   ]
```
This code works because it takes only the values from the row "Albania," thus subsetting by only keeping one country

### **Question 7**:
```{r}
sub3 <- subset(my_data, Population > 100 & Happiness <= 6)
sub3
str(sub3)
```
There are 6 observations in the data set

### **LAB 2**
```{r}
my_data <- read.csv("https://remiller1450.github.io/data/HappyPlanet.csv")
life <- my_data$LifeExpectancy
mean(life)
sd(life)
min(life)
median(life)
max(life)
range(life)
quantile(life, probs = 0.35)
summary(life)
```
### **Question 8**
I like Apples $H_0: \mu = 0$

### **Question 9 P1**
The function range() returns numeric values. The difference between the two numbers is the distance between the min and max

### **Question 9 P2**
```{r}
quantile(my_data$Population, probs = 0.75)-
quantile(my_data$Population, probs = 0.25)
```
```{r}
table(my_data$Region)
table(my_data$LifeExpectancy > 75)
table(my_data$Region, my_data$LifeExpectancy > 75)
tab <- table(my_data$Region, my_data$LifeExpectancy > 75)
class(tab)
prop.table(tab, margin = 1)
prop.table(tab, margin = 2)
```
### **Question 10**
These values are different because for margin = 1, the proportions are compared horizontally, as in the percentages of true and false are measured by region (e.g., 76% of countries in Region 1 have life expectancy over 75). On the other hand, for margin = 2, the proportions are compared vertically, so the percentages are measured by the total proportion of countries that fit under either "true" or "false" (e.g., Region 1 makes up 16% of the countries that have a life expectancy over 75).

```{r}
boxplot(my_data$LifeExpectancy)
tab <- table(my_data$Region)
barplot(tab)
plot(x=women$height, women$weight)
```

### **Question 11 P1**
```{r}
?barplot
barplot(my_data$Region, horiz = TRUE)
```

### **Question 11 P2**
```{r}
?faithful
plot(x=faithful$eruptions, faithful$waiting)
```

It seems as though most of the eruptions either happen after around 1 or 2 minutes, or after 4 or 5 minutes, with the 2.5 to 3.5 minute mark being less frequent. Furthermore, the waiting time to the next eruption increases as the eruption time increases as well.

```{r}
mix_vec <- c(1, 2, 3, "a", "b", "c")
class(mix_vec)
mix_vec <- c(TRUE, FALSE, 1, 2)
mix_vec
class(mix_vec)
mix_vec <- c(TRUE, FALSE, "a")
mix_vec
vec <- c("1", "2", "3")
as.numeric(vec)
mix_vec <- c("1", "2", "3", "a")
as.numeric(mix_vec)
?as.logical

v <- c(T, T, F)

v_num <- as.numeric(v)
v_char <- as.character(v_num)
v_logical <- as.logical(v_num)

v_logical
```

### **Question 12**
as.character makes the "1,1,0" numbers from v_num into characters, but as.logical only regards "True" or "False" character vectors as valid and therefore produces a NA as output. Since numerical vectors are different and can be converted directly into "True" or "False' logical vectors, as.logical works in this case.

```{r}
x <- c ("1","2","3","a")
x <- as.numeric(x)
is.na(x)
!is.na(x)
x[!is.na(x)]
sum(x)
mean(x)
sum(x, na.rm = TRUE)
mean(x, na.rm = TRUE)
```

```{r}
median(my_data$GDPperCapita, na.rm = TRUE)
subt <- subset(my_data, is.na(GDPperCapita))
subt$Country

```
### **Question 13**
The median GDP per capita is 6632. Serbia and Iraq were removed since their GDP per capita wasn't reported.

```{r}
x1 <- c("Dec", "Apr", "Jan", "Mar")
x2 <- c("Dec", "Apr", "Jam", "Mar")
sort(x1)
month_levels <- c("Jan","Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
y1 <- factor(x1, levels = month_levels)
y1
sort(y1)
y2 <- factor(x2, levels = month_levels)
y2
?month.abb
x <- month.abb[c(1,3,2,5,12,11,10,10,3,4,1,3,2,7)]
barplot(table(x))
y <- factor(x, levels = month_levels)
barplot(table(y))
```
```{r}
colleges <- read.csv("https://remiller1450.github.io/data/Colleges2019.csv")
```

### **Question 14**
```{r}
sub1 <- subset(colleges, Adm_Rate < 0.5 & Region == "Great Lakes")
mean(sub1$Salary10yr_median, na.rm = TRUE)
sum(is.na(sub1$Salary10yr_median))

state_levels <- c("IL", "IN", "MI", "OH", "WI")
z1 <- factor(sub1$State, levels = state_levels)
barplot(table(z1))
```

